---
title: "Demo of the pESCA model on Gaussian-Bernoulli-Bernoulli data sets"
author: "Yipeng Song"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GBB_case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This doc is going to show how to simulate Gaussian-Bernoulli-Bernoulli (G-B-B) data sets with underlying global, local common and distinct structures according to the ESCA model. After that, these data sets are used to illustrate how to construct a pESCA model and the corresponding model selection process. In this doc, we will use the simulated parameters to
evluate the model selection process. The documents of the used functions can be found by using the command '? function name'. Since I did't find an easy way to generate the figures in the same way as Matlab, figures will not included in this version.

## Load required packages
```{r load required packages, eval=FALSE}
library(RSpectra)
library(RpESCA)

```
 
## How to simulate proper G-B-B data sets
G-B-B data sets are simulated according to the ESCA model. The number of
samples is set to 200; the number of the variables in the three data sets 
are 1000, 500 and 100; the SNRs in simulating the global, local common and
distinct structures are set to 1; the noise levels (the square root of the 
dispersion parameter $\alpha$) are set to 1; the sparse ratio, which is 
the proportion of 0s in the simulated loading matrix, is set to 0. 

```{r characteristics of simulated data sets, eval=FALSE}
set.seed(123)

# simulate proper data sets
n <- 200
ds <- c(1000, 500, 100)
simulatedData <- dataSimu_group_sparse(n=n,ds=ds,
                                  dataTypes= 'GBB',
                                  noises=rep(1,3),
                                  margProb=0.1,sparse_ratio=0,
                                  SNRgc=1,SNRlc=rep(1,3),SNRd=rep(1,3))

# variation explained ratios for each data set or the full data set
simulatedData$varExpTotals_simu

# variation explained ratios of each PC for each data set or the full data set
simulatedData$varExpPCs_simu

```


## How to estimate the dispersion parameter $\alpha$ for each data set
The dispersion parameters are estimated using the PCA model.
Details can be found in the supplementary information of the paper.

```{r estimate the dispersion parameter, eval=FALSE}
dataSets <- simulatedData$X
dataTypes <- simulatedData$dataTypes

# used parameters
opts <- list()
opts$tol_obj <- 1E-4
opts$quiet <- 1

# save the estimated alphas, selected number of PCs and cvErrors
alphas <- rep(NA,3)
R_selected_list <- as.list(1:3)
cvErrors_list <- as.list(1:3)

# alpha estimation for each data set
for (i in 1:length(dataSets)){
  if(dataTypes[i] == 'G'){
    # index ith data set
  X <- dataSets[[i]]
  
  # alpha estimation procedure
  alpha_est <- alpha_estimation(X,K=3,Rs = 5:20,opts=opts)
  alphas[i] <- alpha_est$alphas_mean
  R_selected_list[[i]] <- alpha_est$R_CV
  cvErrors_list[[i]] <- alpha_est$cvErrors
  }else{
    alphas[i] = 1
  }
}
names(alphas) <- paste0('alpha', 1:3)

# estimated alphas
alphas

# selected number of PCs for the 1-th data set
R_selected_list[[1]]

# cvErrors for the model selection of the 1-th data set
cvErrors_list[[1]]
```

## How to set the parameters for the pESCA model
The following parameters are used for the pESCA model selection. 
The meaning of these parameters can be found in the document of the function.

```{r set up the pESCA model, eval=FALSE}
# concave function and its hyper-parameter
fun_concave <- 'gdp'; gamma <- 1;
penalty = 'L2' # concave L2 norm penalty

# parameters of a pESCA with concave L2norm penalty model
opts <- list()
opts$gamma <- gamma  # hyper-parameter for the used penalty
opts$rand_start <- 0
opts$tol_obj <- 1e-6 # stopping criteria
opts$maxit   <- 500
opts$alphas  <- alphas
opts$R    <- 50    # components used
opts$thr_path <- 0 # generaint thresholding path or not
opts$quiet <- 1

# set up the pESCA model
dataSets <- simulatedData$X
dataTypes <-'GBB'     

```

## The model selection process of the pESCA model
At first, we need to find the proper searching ranges for tuning
$\lambda_g$ for Gaussian data sets and $\lambda_b$ for Bernoulli data
sets. After that, we first fix $\lambda_g$ to a small value and tune
$\lambda_b$, then we will fix $\lambda_b$ and tune $\lambda_g$. 

```{r model selection, eval=FALSE}
# model selection pESCA conave L2 norm penalty
nTries <- 15
lambdas_g <- log10_seq(from=1, to=500, length.out=nTries)
lambdas_b <- log10_seq(from=5, to=100, length.out=nTries)
penalty='L2'

result_CV <- pESCA_CV_twoSteps(dataSets, dataTypes,
                             lambdas_g, lambdas_b,
                             penalty=penalty, fun_concave='gdp',
                             opts=opts)

# CV error changes during the selection of Gaussian data sets
result_CV$cvErrors_g

# CV error changes during the selection of Bernoulli data sets
result_CV$cvErrors_b

# selected value of lambda
result_CV$lambdas_opt

```

## How to fit the final model
After selecting the model with the minimum CV error, the selected model is
re-fitted on the full data set with the selected value of $\lambda$ and
the outputs of the selected model as the initialization. The RV
coefficients in estimating the global common (C123), local common (C12,
C13, C23) and distinct structures (D1, D2, D3) are shown as RVs_structures. 
And the corresponding rank estimations of these structures are shown as 
Ranks_structures. The RMSEs in estimating the simulated
$\mathbf{\Theta}$, $\mathbf{\Theta}_1$, $\mathbf{\Theta}_2$ and
$\mathbf{\Theta}_3$ and $\mu$ are shown as RMSEs_parameters. 

```{r fit the final model, eval=FALSE}
# fit the final model
lambdas_opt <- result_CV$lambdas_opt
opts_opt <- result_CV$opts_opt
opts_opt$tol_obj <- 1E-8 # using high precision model

pESCA_L2 <- pESCA(dataSets = dataSets,
                          dataTypes = dataTypes,
                          lambdas = lambdas_opt,
                          penalty = penalty,
                          fun_concave= fun_concave,
                          opts = opts_opt)

# evaluate the final model
mu <- pESCA_L2$mu
A <- pESCA_L2$A
B <- pESCA_L2$B
S <- pESCA_L2$S
pESCA_L2_eval <- eval_metrics_simu_group(mu,A,B,S,ds,simulatedData)

# RV coefficients in estimating the simulated structures
pESCA_L2_eval$RVs_structs

# Rank estimations of the simulated structures
pESCA_L2_eval$Ranks_structs

# RMSEs in estimating the simulated parameters
pESCA_L2_eval$RMSEs_params

# estimated variation explained ratios for each data set or the full data set
pESCA_L2$varExpTotals

# estimated variation explained ratios of each PC for each data set or the full data set
pESCA_L2$varExpPCs

```


